---
title: "Workbook 1: Set Retrieval Metrics"
author: "Maximilian Kähler, DNB"
format: html
editor: source
---

```{r}
#| label: setup
#| message: false
#| warning: false
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(knitr)
library(tidyverse)
library(casimir)
options(casimir.ignore_inconsistencies = TRUE)
```

```{r}
#| label: read-data
#| echo: FALSE


gold_standard <- read_csv("../data/test-set_gold-standard.csv",
                          col_select = c("doc_id", "label_id"))

files <- list(
  "method-A" = "../data/test-set_predictions/method-A.csv",
  "method-B" = "../data/test-set_predictions/method-B.csv",
  "method-C" = "../data/test-set_predictions/method-C.csv",
  "method-D" = "../data/test-set_predictions/method-D.csv",
  "method-E" = "../data/test-set_predictions/method-E.csv",
  "method-F" = "../data/test-set_predictions/method-F.csv"
)
predictions <- files |> 
  map(read_csv) 
```

# Set Retrieval Metrics

The basic way to compute metrics with casimir is the function 
`compute_set_retrieval_scores()`. It takes as input a set of predicted labels 
and a gold standard set of labels and computes various retrieval metrics such 
as precision, recall, and F1-score. 

Short reminder: Precision is the fraction of suggested subject terms that is
correct, while recall is the fraction of gold standard subject terms that are
retrieved. F1-score is the harmonic mean of precision and recall. 

R-Precision is the precision at R, where R is the number of gold standard 
subjects. This sort of avoids penelizing methods that suggest way too many
subjects, by limiting to the actual amount of relevant subjects that could be
found.

The parameter `k` allows you to specify how many of the top predicted labels
should be considered for the computation of the metrics.

```{r}
#| label: compute-metrics

compute_set_retrieval_scores(
  predicted = predictions[["method-A"]],
  gold_standard = gold_standard,
  k = 5,
  rename_metrics = TRUE
)

```

You can iterate that over multiple methods using R's `map_*()` functions for functional
programming. `map_dfr()` is particularly useful here as it combines the results 
into a single data frame.

```{r}
#| label: compute-metrics-multiple

results <- map_dfr(
  predictions,
  ~ compute_set_retrieval_scores(
    predicted = .x,
    gold_standard = gold_standard,
    k = 100,
    rename_metrics = TRUE
  ),
  .id = "method"
)

kable(results)
```

Let's make a first visualization using R`s ggplot package

```{r}
#| label: plot-metrics

ggplot(results, aes(x = method, y = value, fill = method)) +
  geom_bar(stat = "identity") +
  ylim(0, 1) +
  facet_wrap(~ metric) +
  theme_minimal() +
  labs(
    title = "Retrieval Metrics by Method",
    x = "Method",
    y = "Score"
  ) + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```


# Your Turn

Play around with the input options for `compute_set_retrieval_scores()` (`?compute_set_retrieval_scores` is your friend)

  * Observe how you can increase precision by lowering the `k` parameter, but at the cost of recall. 
  * What is the theoretical recall limit that you achieve by setting `k = 100` ?
  * what happens when you switch aggregation modes `mode = "micro"` vs `mode = "doc-avg"` vs. `mode = "subj-avg"`? Can you explain the differences?
  * what are your preffered methods, so far?