---
title: "Workbook 2: Stratified Set Retrieval Metrics"
author: "Maximilian Kähler, DNB"
format: gfm
toc: true
editor: source
---

We are now going to learn how to compute stratified set retrieval metrics. This
is an important technique to evaluate the performance of subject indexing systems
on different subsets of data, such as documents belonging to different categories
or subjects.

```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: FALSE
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.path = "figures/02_stratified-set-retrieval-metrics/"
  )

library(knitr)
library(tidyverse)
library(casimir)
options(casimir.ignore_inconsistencies = TRUE, digits = 3)
```

```{r}
#| label: read-data
#| echo: FALSE


gold_standard <- read_csv("../data/test-set_gold-standard.csv",
                          col_select = c("doc_id", "label_id"),
                          show_col_types = FALSE)

files <- list(
  "method-A" = "../data/test-set_predictions/method-A.csv",
  "method-B" = "../data/test-set_predictions/method-B.csv",
  "method-C" = "../data/test-set_predictions/method-C.csv",
  "method-D" = "../data/test-set_predictions/method-D.csv",
  "method-E" = "../data/test-set_predictions/method-E.csv",
  "method-F" = "../data/test-set_predictions/method-F.csv"
)
predictions <- files |> 
  map(read_csv, show_col_types = FALSE) 

subject_groups <- read_csv("../data/test-set_subject-group-mapping.csv",
                           col_select = c(
                            "doc_id", "sg", "sg_label_ger", "sg_label_eng"),
                            show_col_types = FALSE)


gnd_entity_types <- read_csv("../data/gnd_entitytypes.csv",
                             col_select = c(
                              "label_id", 
                              "label_entitytype_ger",
                              "label_entitytype_eng"
                              ),
                             show_col_types = FALSE
                            )
```

## Stratified results by subject groups

In the next code chunk we compute the set retrieval scores at rank 5 for
documents stratified by their subject groups using the `compute_set_retrieval_scores()` function. This is as simple as
the `doc_groups` argument with a prepared data frame containing the document IDs and their corresponding subject groups.

```{r}
#| label: stratified-subject-groups
res_at_5_by_sg_method_A <- compute_set_retrieval_scores(
  predicted = predictions[["method-A"]],
  gold_standard = gold_standard,
  doc_groups = subject_groups,
  k = 5
)

head(res_at_5_by_sg_method_A)

```

**Note on terminology**: We refer to "subject groups" here as groups of documents that share a common subject classification, e.g. "History" or "Science". This is different from "subjects" in the sense of "subject terms" or "subject headings", which are the actual labels assigned to documents.

As the information contained in the resulting data frame is quite extensive, we 
apply some post-processing to simplify the display of results. We filter for
the metrics precision, recall, and F1-score, remove unnecessary columns, and
pivot the data frame to a wider format for better readability.

```{r}
#| label: stratified-subject-groups-display
# apply post processing to simplify display of results
res_at_5_by_sg_method_A  |>
  filter(metric %in% c("prec", "rec", "f1")) |> 
  select(-support, -mode)  |>
  pivot_wider(
    names_from = metric,
    values_from = value
  )  |>
  kable(caption = "Set retrieval scores at rank 5 for Method A stratified by subject groups.")
```

We can see that the performance of Method A varies across different subject 
groups. For example, the f1-score for physics is best, whereas
computer science has the lowest f1-score.

While this sort of analysis can be done manually for each method, it is often more
convenient to compute the stratified results for all methods in one go. 

```{r}
#| label: stratified-subject-groups-all-methods
res_at_5_by_sg_all_methods <- map_dfr(
  predictions,
  ~ compute_set_retrieval_scores(
    predicted = .x,
    gold_standard = gold_standard,
    doc_groups = subject_groups,
    k = 5),
  .id = "Method"
)

head(res_at_5_by_sg_all_methods)
```

As above, we apply some post-processing to simplify the display of results.

```{r}
#| label: stratified-subject-groups-all-methods-display
# apply post processing to simplify display of results
res_at_5_by_sg_all_methods  |>
  filter(metric == "f1") |>
  select(-support, -mode)  |>
  pivot_wider(
    names_from = Method,
    values_from = value
  )  |>
  kable(caption = "F1-scores at rank 5 for all methods stratified by subject groups.")
```

Here we have reached a level of detail, where displaying only numbers makes it 
quite challenging to interpret the results. Visualizations can help to better understand
the performance of different methods across subject groups.

```{r}
#| label: stratified-subject-groups-plot
#| fig.width: 10
#| fig.height: 20
res_at_5_by_sg_all_methods  |>
  filter(metric == "f1") |>
  ggplot(aes(x = Method, y = value, fill = Method)) +
  ylim(0, 0.7) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(
    ~sg_label_eng,
    ncol = 4,
    labeller = label_wrap_gen(width = 20)
  ) +
  labs(
    title = "F1-scores at rank 5 for all methods stratified by subject groups",
    x = "Method",
    y = "F1-score"
  ) + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```