---
title: "Workbook 2: Set Retrieval Metrics"
author: "Maximilian Kähler, DNB"
format: gfm
toc: true
editor: source
---

In this workbook we will learn how to compute the basic set retrieval metrics
like precision, recall, and F1-score using the CASIMiR package.
We will also explore how to drill down into results on the subject level
to find the best and worst performing subject terms.

```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: FALSE
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.path = "figures/02_set-retrieval-metrics/"
  )

library(knitr)
library(tidyverse)
library(casimir)
options(casimir.ignore_inconsistencies = TRUE)
```

```{r}
#| label: read-data
#| echo: FALSE

source("../src/load_data.r")

```

## Computing overall set retrieval metrics

The basic way to compute metrics with CASIMIiR is the function 
`compute_set_retrieval_scores()`. It takes as input a set of predicted labels 
and a gold standard set of labels and computes various retrieval metrics such 
as precision, recall, and F1-score. 

::: {.panel-tabset}
#### More about metrics

<details>
<summary>Click to expand explanation of metrics</summary>

Short reminder: **precision** is the fraction of suggested subject terms that is
correct, while **recall** is the fraction of gold standard subject terms that are
retrieved. Thus, while the denominator for recall is based on the gold standard,
the denominator for precision is based on the number of predictions you 
choose to make. 

**F1-score** is the harmonic mean of precision and recall. 

**R-Precision** is the precision at $R$, where $R$ is the number of gold standard 
subjects. This sort of avoids penalizing methods that suggest way too many
subjects, by limiting predictions to the actual amount of relevant subjects that could be found.

</details>
:::

The parameter `k` allows you to specify how many of the top predicted labels
should be considered for the computation of the metrics.

```{r}
#| label: compute-metrics-one-method

compute_set_retrieval_scores(
  predicted = predictions[["artful-accordion"]],
  gold_standard = gold_standard,
  k = 5,
  rename_metrics = TRUE
)

```

The column `support` indicates how many documents have contributed to the
computation of the respective metric. If `artful-accordion` made predictions for
all documents in the gold standard, this should be equal to the number of
documents in the gold standard. But if some documents were not assigned any
predicted labels, they are not considered for the computation of precision.
See below digression for more details on aggregation modes and zero division
handling.

You can iterate the function `compute_set_retrieval_scores()`
over multiple methods using the `map_*()` functions from the
tidyverse `purrr` package for functional data processing. 
`map_dfr()` is particularly useful here as it combines the results 
into a single data frame.

```{r}
#| label: compute-metrics-at-K-multiple
K <- 5

results <- map_dfr(
  predictions,
  ~ compute_set_retrieval_scores(
    predicted = .x,
    gold_standard = gold_standard,
    k = K,
    rename_metrics = TRUE
  ),
  # create id column from list names
  .id = "Method"
)

# bring results to wide tible for better display
results |> 
  select(-support) |> 
  pivot_wider(
    names_from = metric,
    values_from = value
  ) |>
kable(caption = paste0("Set retrieval metrics for all methods (k = ", K, ")"))

```

Let's make a first visualization using R`s ggplot package

```{r}
#| label: plot-metrics

ggplot(results, aes(x = Method, y = value, fill = Method)) +
  geom_bar(stat = "identity") +
  ylim(0, 1) +
  facet_wrap(~ metric) +
  theme_minimal() +
  labs(
    title = "Set Retrieval Metrics by Method",
    x = "Method",
    y = "Score"
  ) + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

<details>
<summary>Note</summary>
Some people prefer to "zoom-in" on a smaller scale of the y-range to
emphasize the differences between methods. While this is sometimes acceptable 
for visualization purposes, be aware that this can be misleading as it visually
exaggerates small differences. Always check the actual values in the data
table to get a true sense of the performance differences.
</details>


## Finding buttom and top performerming labels  

If you want to drill down into results on subject level, CASIMiR also provides 
functions to disect results on the intermediate aggregation level. 
The above function `compute_set_retrieval_scores()` is just a convenient 
wrapper around three steps:

  1. `create_comparison()`: creates a detailed comparison of predicted vs. 
    gold standard labels for each doc_id, label_id pair.
  2. `compute_intermediate_results()`: computes true positives, false positives, 
    and false negatives on the desired aggregation level (per document or per 
    subject).
  3. `summarise_intermediate_results()` computes avarages of the intermediate 
    results to yield the final precision, recall, and F1-score.
    
Let's use this to find out the best and worst performing subject terms for
`artful-accordion`.
    
```{r}
#| label: Top-and-bottom-labels

label_texts <- read_csv("../data/gnd_pref-labels_w-translation.csv",
                        col_select = c("label_id", "label_text_eng"))

comp <- create_comparison(
  predicted = predictions[["artful-accordion"]],
  gold_standard = gold_standard
)

intermed <- compute_intermediate_results(
  gold_vs_pred = comp,
  grouping_var = "label_id"
)$results_table |> 
  left_join(label_texts, by = "label_id")

intermed |>
  filter(n_gold > 20) |> 
  arrange(desc(f1)) |> 
  select(label_id, label_text_eng, tp, fp, fn, prec, rec, f1) |> 
  head(n = 10) |> 
  kable(
    caption = "Best performing subject terms for artful-accordion with a mininimum of
     20 gold standard instances."
  )

intermed |>
  filter(n_gold > 20) |> 
  arrange(f1) |> 
  select(label_id, label_text_eng, tp, fp, fn, prec, rec, f1) |> 
  head(n = 10) |> 
  kable(
    caption = "Worst performing subject terms for artful-accordion with a minimum of
     20 gold standard instances."
  )
  
```

## Your Turn

Play around with the input options for `compute_set_retrieval_scores()` 
(`?compute_set_retrieval_scores` is your friend)

  * Observe how you can increase precision by lowering the `k` parameter, but
    at the cost of recall. 
  * What is the theoretical recall limit that you achieve by setting `k = 100` ?
  * compare the top and bottom performing labels for other methods
  * what happens when you switch aggregation modes `mode = "micro"` vs `mode = "doc-avg"` vs. `mode = "subj-avg"`? 
    Can you explain the differences? (see below for more details)
  * what are your preferred methods, so far?
  
## Bonus: A short digression to modes of aggregation

The `compute_set_retrieval_scores()` function supports three modes of aggregation

  * `mode = "doc-avg"`: (default) computes precision, recall, and F1-score for each 
    **document** individually and then average these scores across all documents.
  * `mode = "subj-avg"`: computes precision, recall, and F1-score for each 
    **subject term** individually and then average these scores across all subject 
    terms.
  * `mode = "micro"`: computes global counts of true positives, false positives, 
    and false negatives across all documents and subjects, and then calculates
    precision, recall, and F1-score based on these aggregated counts.

`subj-avg` and `doc-avg` are two flavours of so called macro-averaging, where
you first compute metrics on a more granular level (per document or per subject)
and then average them. This way, each document or subject contributes equally to
the final score, regardless of how many predictions were made for each.
In contrast, `micro` aggregation pools all predictions together and computes
overall counts. This way, documents or subjects with more predictions have a
larger impact on the final score.

The column `support` indicates how many instances on the intermediate level
have contributed to the final score. For `doc_id` grouping, this is the number
of documents. For `label_id` grouping, this is the number of subject terms.
By default, CASIMiR ignores documents or subject terms that have no predictions
when computing precision macro averages. This avoids zero-division issues.
To be coherent with other packages like [scikit-learn](https://scikit-learn.org/stable/modules/generated/sklearn.metrics.f1_score.html), you can change this behaviour by using the 
argument `replace_zero_division_with = 0` which will set precision to 0
for documents or subject terms with no predictions, if they have occurences in 
the gold standard. 

Observe how much this can influence the overall value of precision and 
recall in subject averaging mode:

```{r}
#| label: subj-avg-zero-devision
compute_set_retrieval_scores(
  predicted = predictions[["artful-accordion"]],
  gold_standard = gold_standard,
  k = 5,
  mode = "subj-avg",
  rename_metrics = TRUE,
  replace_zero_division_with = 0 # set to NA for default behaviour
)
```
