---
title: "Workbook 3: Precision Recall Curves"
author: "Maximilian Kähler, DNB"
format: gfm
toc: true
editor: source
---

So far we have studied set retrieval metrics: metrics that assume a fixed
set of predictions. In most exercises we have already used that, in practice, 
many indexing methods produce ranked lists of candidates. We previously applied
a limit `k = 5`, telling CASIMiR to only look at the top-5 predictions per 
document. These ranks are based on **confidence scores** for each
suggested subject candidate. Instead of cutting off at a fixed rank `k`, we can also
cut off at a fixed confidence score threshold. This provides another way to vary
the trade-off between **precision** (how many of the suggested subjects are correct) 
and **recall** (how many of the correct subjects are suggested). 
In this exercise, we will compute precision-recall curves, visualizing this 
trade-off for different for different methods.


```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: FALSE
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.path = "figures/03_precision-recall-curves/"
  )

library(knitr)
library(tidyverse)
library(future)
library(casimir)
options(casimir.ignore_inconsistencies = TRUE, digits = 3)
```


```{r}
#| label: read-data
#| echo: FALSE

source("../src/load_data.r")

```

## The idea of precision-recall curves

To get to know precision-recall curves, let's first do all the computations
step-by-step, before we later use CASIMiR's built-in functions to compute
precision-recall curves directly.

Let's define a function that computes precision and recall for method `artful-accordion`
at **varying limits** `k`:

```{r}
#| label: define-prec-rec-rank

# compute precision and recall at k for method artful-accordion
prec_rec_rank <- function(k) {
  res_at_k <- compute_set_retrieval_scores(
    predicted = predictions[["artful-accordion"]],
    gold_standard = gold_standard,
    k = k
  )  

  # simplify output to only precision and recall
  res <- res_at_k  |>
    filter(metric %in% c("prec", "rec"))  |>
    select(metric, value)  |>
    pivot_wider(names_from = metric, values_from = value)  |>
    select(prec, rec)  |>
    mutate(k = k)

  res
}

prec_rec_rank(5)
```

Now we can compute precision and recall for all limits `k` from 1 to 20 and
plot the results:
```{r}
#| label: plot-prec-rec-rank
#| cache: true

k_range <- 1:20
prec_rec_rank_df <- map_dfr(k_range, prec_rec_rank)

ggplot(prec_rec_rank_df, aes(x = rec, y = prec)) +
  geom_path() +
  geom_point() +
  geom_text(aes(label = paste0("k = ", k)), vjust = -1) +
  coord_fixed(xlim = c(0, 1), ylim = c(0, 1)) +
  labs(
    title = "Precision-Recall Curve for artful-accordion (by rank)",
    x = "Recall",
    y = "Precision"
  )

```

The same idea can be applied not to the **rank** but to the **confidence score**.
Let's define a function that computes precision and recall for method `artful-accordion`
at **varying confidence score thresholds**:
```{r}
#| label: define-prec_rec_threshold

prec_rec_threshold <- function(threshold) {
  # filter predictions by threshold
  predictions_thresholded <- predictions[["artful-accordion"]] |>
    filter(score >= threshold)
  
  # compute precision and recall at threshold
  res_at_threshold <- compute_set_retrieval_scores(
    predicted = predictions_thresholded,
    gold_standard = gold_standard,
  )  

  # simplify output to only precision and recall
  res <- res_at_threshold  |>
    filter(metric %in% c("prec", "rec"))  |>
    select(metric, value)  |>
    pivot_wider(names_from = metric, values_from = value)  |>
    select(prec, rec)  |>
    mutate(threshold = threshold)

  res
}

prec_rec_threshold(0.1)
```

Again, we can compute precision and recall for a range of thresholds from 0 to 1
and plot the results:
```{r}
#| label: plot-prec-rec-threshold
#| cache: true

threshold_range <- seq(0, 1, by = 0.1)
prec_rec_threshold_df <- map_dfr(threshold_range, prec_rec_threshold)

ggplot(prec_rec_threshold_df, aes(x = rec, y = prec)) +
  geom_path() +
  geom_point() +
  geom_text(aes(label = paste0("t = ", round(threshold, 2))), hjust = -0.5) +
  coord_fixed(xlim = c(0, 1), ylim = c(0, 1)) + 
  labs(
    title = "Precision-Recall Curve for artful-accordion (by threshold)",
    x = "Recall",
    y = "Precision"
  )
 

```

If you increase the number of steps, e.g. `threshold_range <- seq(0, 1, by = 0.05)`,
you may note some "wiggeling" (non-monotonicity) of the pr-curve. 
Increasing the threshold does not always lead to an increase in precision. 
In this case one would, in practice, use the lower threshold that achieves
higher recall with higher precision. This is why usually precision-recall curves
do not plot the precision
but the **cumulative maximum precision** (`prec_cummax`), which is guaranteed
to be non-increasing with increasing recall. `prec_cummax` at recall level `r` 
is defined as the maximum precision achieved at any recall `r' >= r`.

## Precision-recall curves with CASIMiR

It is tedious and error-prone to compute precision-recall curves
step-by-step as we did above. CASIMiR provides a convenient function
to compute precision-recall curves directly:
```{r}
#| label: prec-rec-casimir
#| cache: true

pr_curve <- compute_pr_curve(
  predicted = predictions[["artful-accordion"]],
  gold_standard = gold_standard,
  steps = 10
)

ggplot(pr_curve$plot_data, aes(x = rec, y = prec_cummax)) +
  geom_point() +
  geom_path() +
  coord_fixed(xlim = c(0, 1), ylim = c(0, 1)) + 
  ggtitle("Precision-Recall Curve for artful-accordion, computed with CASIMiR")
```

By default, CASIMiR computes the pr-curve using thresholding over confidence
scores. The threshold range is controlled by the `steps` argument, which defaults
to 100. Here we work with a lower number of steps for faster computation.
You can also include a search over ranks by adding the argument
`limit_range = 1:5` or similar. This will compute precision and recall for all
combinations of thresholds and limits in the specified range, and find the highest
precision that can be attained at a given recall value.

## Area under the precision recall curve (pr-AUC)

PR-Curves are interesting to study in its own right, but it is often useful
to summarize the curve in a single number. The most common summary statistic
is the **area under the curve** (AUC or pr-AUC). This is simply the area below the
precision-recall curve. The intuition is, that a curve that comes close to the 
upper right corner (high precision and high recall) will have a high area below it,
while a curve that is close to the lower left corner (low precision and low recall)
will have a low area below it.

CASIMiR offers a function to compute the pr-AUC directly from the pr-curve data:
```{r}
#| label: compute-auc 
pr_auc <- compute_pr_auc_from_curve(pr_curve)

pr_auc
```

Alternatively, the AUC can also be computed directly from predictions and
gold standard without computing the full pr-curve first:
```{r}
#| label: compute-auc-direct
#| cache: true
pr_auc_direct <- compute_pr_auc(
  predicted = predictions[["artful-accordion"]],
  gold_standard = gold_standard,
  steps = 10
)

pr_auc_direct

```

`compute_pr_auc` accepts almost all arguments that `compute_set_retrieval_scores`
accepts, e.g. `subject_groups` and `label_groups`, allowing to compute AUC
for specific strata of the data.

### Finding optimal cutoffs

Often it is not useful to know the theoretical recall at a very low precision,
or vice versa. In production systems, one often wants to find the sweet spot
between precision and recall that fits the use case best. CASIMiR provides a
function to find such optimal cutoffs based on optimal F1-score:
```{r}
#| label: find-optimal-cutoff
#| cache: true

# optionally use parallel processing for faster computation
library(future)
plan(multicore)

pr_curve_with_cutoff <- compute_pr_curve(
  predicted = predictions[["artful-accordion"]],
  gold_standard = gold_standard, 
  steps = 10,
  limit_range = 1:10,
  optimize_cutoff = TRUE,
  progress = TRUE
)

plan(sequential) # reset to sequential processing

ggplot(pr_curve_with_cutoff$plot_data, aes(x = rec, y = prec_cummax)) +
  geom_point(
    data = pr_curve_with_cutoff$opt_cutoff,
    aes(x = rec, y = prec_cummax),
    color = "red",
    shape = "star"
  ) +
  geom_text(
    data = pr_curve_with_cutoff$opt_cutoff,
    aes(
      x = rec + 0.2, y = prec_cummax,
      label = paste("f1_opt =", round(f1_max, 3))
    ),
    color = "red"
  ) +
  geom_path() +
  coord_fixed(xlim = c(0, 1), ylim = c(0, 1)) + 
  labs(
    title = "Precision-Recall Curve with optimal F1-score",
    x = "Recall",
    y = "Precision"
  )
```

CASIMiR also tells you at what limit and threshold the optimnal F1-score
is achieved:
```{r}
#| label: show-optimal-cutoff
pr_curve_with_cutoff$opt_cutoff
```


::: {.panel-tabset}
#### Digression: In-Sample BIAS

<details>
<summary>Click to expand for information on in-sample bias</summary>


**Warning:** When optimizing cutoffs on the same data that is used to
evaluate the performance, there is a risk of **in-sample bias**.
This means that the optimal cutoff found may be overly optimistic
for the given data, and may not generalise well to new, unseen data.
One can mitigate this risk by using multiple test-sets or cross-validation,
where the optimal cutoff is determined on one test set and then
evaluated on a separate second test set.

</details>
:::

## Comparing PR-Curves of multiple methods

Finally, we can compare the PR-curves of multiple methods:

```{r}
#| label: compute-multiple-pr-curves
#| cache: true

plan(multicore)

pr_curves <- map(
  predictions,
  ~ compute_pr_curve(
      predicted = .x,
      gold_standard = gold_standard,
      limit_range = 1:10,
      optimize_cutoff = TRUE,
      progress = TRUE
    )
)

plan(sequential)

```

```{r}
#| label: plot-multiple-pr-curves

# extract curves
all_curves <- imap_dfr(
  pr_curves,
  ~ mutate(.x$plot_data, method = .y)
)

# extract optimal f1 scores
f1_opt <- imap_dfr(
  pr_curves,
  ~ mutate(.x$opt_cutoff, method = .y)
)

# compute area under the curve (AUC) for all methods
pr_auc <- map_dfr(
  pr_curves,
  compute_pr_auc_from_curve,
  .id = "method"
)

ggplot(all_curves, aes(x = rec, y = prec_cummax, color = method)) +
  geom_point(
    data = f1_opt,
    aes(x = rec, y = prec_cummax),
    shape = "star"
  ) +
  geom_text(
    data = f1_opt,
    aes(
      x = rec + 0.2, y = prec_cummax,
      label = paste("f1_opt =", round(f1_max, 3))
    )
  ) +
  geom_path() +
  coord_fixed(xlim = c(0, 1), ylim = c(0, 1))

f1_opt |>
  select(method, f1_max)  |>
  kable(caption = "Optimal F1-scores for all methods")

pr_auc |>
  kable(caption = "Area under the precision-recall curve (pr-AUC) for all methods")
```

## Your turn

  * what are your observations from the PR-curves of the different methods?
  * which method scores best, second best, etc. in terms of pr-AUC and optimal F1-score? Do the rankings differ?
  * go back to workbook 1 and compare your findings with the set retrieval metrics at fixed k=5. Are the rankings similar or different?

## Bonus exercise: Fusing predictions

Let's think about the theoretical preformance that we can achieve simply 
by combining all methods.

The following code generates a fused list of predictions by summing up the 
confidence scores of all methods and normalizing them to the range [0,1]:

```{r}
#| label: all-in-one
#| cache: true

# define fusion strategy
fuse_predictions <- function(x, y) {

  full_join(x,y, by = c("doc_id", "label_id"))  |>
    mutate(
      score = coalesce(score.x, 0) + coalesce(score.y, 0)
    )  |>
    select(doc_id, label_id, score)
} 

# iteratively fuse all predictions
fused_predictions <- reduce(predictions, .f = fuse_predictions)  |>
  mutate(score = score/length(predictions)) # normalize to [0,1] range
```

```{r}
#| label: fused-pr-curve
#| cache: true

plan(multicore)

fused_pr_curve <- compute_pr_curve(
  predicted = fused_predictions,
  gold_standard = gold_standard,
  limit_range = 1:10,
  optimize_cutoff = TRUE
)

plan(sequential)

```

```{r}
#| label: plot-fused-pr-curve

ggplot(fused_pr_curve$plot_data, aes(x = rec, y = prec_cummax)) +
  geom_point(
    data = fused_pr_curve$opt_cutoff,
    aes(x = rec, y = prec_cummax),
    color = "red",
    shape = "star"
  ) +
  geom_text(
    data = fused_pr_curve$opt_cutoff,
    aes(
      x = rec + 0.2, y = prec_cummax,
      label = paste("f1_opt =", round(f1_max, 3))
    ),
    color = "red"
  ) +
  geom_path() +
  coord_fixed(xlim = c(0, 1), ylim = c(0, 1)) + 
  labs(
    title = "Precision-Recall Curve for Fused Predictions",
    x = "Recall",
    y = "Precision"
  )

pr_auc_fused <- compute_pr_auc_from_curve(fused_pr_curve)

pr_auc_fused
```

This sort of fusion is only the very simplest form of ensemble learning. 
More sophisticated
methods could weight different methods differently, or use machine learning
to learn an optimal combination of methods based on features of the documents
or subjects. The key takeaway here is that by combining multiple methods,
one can often achieve better performance than any individual method alone.