---
title: "Workbook 3: Stratified Set Retrieval Metrics"
author: "Maximilian Kähler, DNB"
format: gfm
toc: true
editor: source
---

We are now going to learn how to compute stratified set retrieval metrics. This
is an important technique to evaluate the performance of subject indexing systems
on different subsets of data, such as documents belonging to different categories
or subjects.

```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: FALSE
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.path = "figures/03_stratified-set-retrieval-metrics/"
  )

library(knitr)
library(tidyverse)
library(casimir)
options(casimir.ignore_inconsistencies = TRUE, digits = 3)
```

```{r}
#| label: read-data
#| echo: FALSE

source("../src/load_data.r")

# load mapping table of documents with their subject groups
subject_groups <- read_csv("../data/test-set_subject-group-mapping.csv",
                           col_select = c(
                            "doc_id", "sg", "sg_label_ger", "sg_label_eng"),
                            show_col_types = FALSE)  |>
                    # combine some subject groups for better readability
                    mutate(
                      sg_label_eng = as.factor(case_when(
                        sg %in% c("940", "943") ~ "History of Germany and Europe",
                        TRUE ~ sg_label_eng
                      )),
                      sg_label_ger = as.factor(case_when(
                        sg %in% c("940", "943") ~ "Geschichte Deutschlands und Europas",
                        TRUE ~ sg_label_ger
                      )),
                      sg = as.factor(case_when(
                        sg %in% c("940", "943") ~ "940/943",
                        TRUE ~ sg
                      ))
                    )
                          

# load mapping table of labels with their entity types
gnd_entity_types <- read_csv("../data/gnd_entitytypes.csv",
                             col_select = c(
                              "label_id", 
                              "label_entitytype_ger",
                              "label_entitytype_eng"
                              ),
                             show_col_types = FALSE
                            )
```

## Stratified results by document groups

In the next code chunk we compute the set retrieval scores at rank 5 for
documents stratified by their subject groups using the `compute_set_retrieval_scores()` function. This is as simple as using
the `doc_groups` argument with a prepared data frame containing the document IDs and their corresponding subject groups.

```{r}
#| label: stratified-subject-groups
res_at_5_by_sg_artful_accordion <- compute_set_retrieval_scores(
  predicted = predictions[["artful-accordion"]],
  gold_standard = gold_standard,
  doc_groups = subject_groups,
  k = 5
)

head(res_at_5_by_sg_artful_accordion)

```

**Note on terminology**: We refer to "subject groups" here as groups of 
documents that share a common subject classification, e.g. "History" or 
"Science". This is different from "subjects" in the sense of "subject terms"
 or "subject headings", which are the actual labels assigned to documents in the
 process of subject indexing.

As the information contained in the resulting data frame is quite extensive, we 
apply some post-processing to simplify the display of results. We filter for
the metrics precision, recall, and F1-score, remove unnecessary columns, and
pivot the data frame to a wider format for better readability.

```{r}
#| label: stratified-subject-groups-display
# apply post processing to simplify display of results
res_at_5_by_sg_artful_accordion  |>
  filter(metric %in% c("prec", "rec", "f1")) |> 
  select(-support, -mode)  |>
  pivot_wider(
    names_from = metric,
    values_from = value
  )  |>
  kable(caption = "Set retrieval scores at rank 5 for artful-accordion stratified by subject group.")
```

We can see that the performance of artful-accordion varies across different subject 
groups. For example, the f1-score for physics is best, whereas
computer science has the lowest f1-score.

While this sort of analysis can be done manually for each method, it is often more
convenient to compute the stratified results for all methods in one go. 

```{r}
#| label: stratified-subject-groups-all-methods

# iterate over all methods and compute set retrieval scores by subject groups
res_at_5_by_sg_all_methods <- map_dfr(
  predictions,
  ~ compute_set_retrieval_scores(
    predicted = .x,
    gold_standard = gold_standard,
    doc_groups = subject_groups,
    k = 5),
  .id = "Method"
)

head(res_at_5_by_sg_all_methods)
```

As above, we apply some post-processing to simplify the display of results.

```{r}
#| label: stratified-subject-groups-all-methods-display
# apply post processing to simplify display of results
res_at_5_by_sg_all_methods  |>
  filter(metric == "f1") |>
  select(-support, -mode)  |>
  pivot_wider(
    names_from = Method,
    values_from = value
  )  |>
  kable(caption = "F1-scores at rank 5 for all methods stratified by subject groups.")
```

Here we have reached a level of detail, where displaying only numbers makes it 
quite challenging to interpret the results. Visualizations can help to better understand
the performance of different methods across subject groups.

```{r}
#| label: stratified-subject-groups-plot
#| fig.width: 10
#| fig.height: 20
res_at_5_by_sg_all_methods  |>
  filter(metric == "f1") |>
  ggplot(aes(x = Method, y = value, fill = Method)) +
  ylim(0, 0.7) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(
    vars(paste0(sg, ": ", sg_label_eng)),
    ncol = 4,
    labeller = label_wrap_gen(width = 20)
  ) +
  labs(
    title = "F1-scores at rank 5 for all methods stratified by subject groups",
    x = "Method",
    y = "F1-score (doc-avg)"
  ) + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Your turn

Some questions that may help to reflect on the results:

  * which methods would win in the respective subject groups? 
  * are there subject groups with similar performance characteristics across 
    methods? 
  * are there subject groups that show very distinct performance characteristics 
    from the other groups? 
  * alter the above plot to show precision or recall instead of f1-score. 
    Do you observe different trends?


## Bonus: Confidence Intervals

The smaller we make the strata in our analysis, the larger will be the 
uncertainty of performance estimates in these strata. We should not jump to
conclusions based on results that are statistically unsound. This is why
it is important to take into account uncertainty by computing confidence 
intervals. Using the option `compute_bootstrap_ci = TRUE` will compute
95% confidence intervals for each score. These are based on the 
[bootstrap method](https://stat20.berkeley.edu/fall-2024/3-generalization/09-bootstrapping/notes.html).
Bootstrapping is a resampling technique that involves repeatedly sampling
with replacement from the original data to create "bootstrap samples". The
metrics of interest (e.g., f1, precision, recall) are calculated for each
bootstrap sample, resulting in a distribution of the metric. This distribution
can then be used to estimate confidence intervals. The parameter `n_bt`
specifies the number of bootstrap samples to generate. The more samples,
the more reliable the confidence intervals, but also the longer the computation time. 
When computing confidence intervals, it is often beneficial to use parallel
processing to speed up the computation. This can be done using the 
[future](https://future.futureverse.org/) package.

```{r}
#| label: stratified-subject-groups-ci
#| cache: true

# optionally use parallel processing for faster computation
library(future)
plan(multicore)

res_at_5_by_sg_artful_accordion_ci <- compute_set_retrieval_scores(
  predicted = predictions[["artful-accordion"]],
  gold_standard = gold_standard,
  doc_groups = subject_groups,
  k = 5,
  compute_bootstrap_ci = TRUE,
  n_bt = 100L, # number of bootstrap samples for confidence intervals
  progress = TRUE
)

head(res_at_5_by_sg_artful_accordion_ci)

plan(sequential) # reset to sequential processing
```

The resulting data frame now contains additional columns for the lower
and upper bounds of the confidence intervals (`ci_lower` and `ci_upper`).
These can be used to visualize the uncertainty of the estimates in a plot:

```{r}
#| label: stratified-subject-groups-ci-display
#| fig.width: 10
#| fig.height: 20

# plot results with confidence intervals
ggplot(
  res_at_5_by_sg_artful_accordion_ci,
  aes(x = sg, y = value, ymin = ci_lower, ymax = ci_upper)
) + 
  geom_pointrange(position = position_dodge(width = 0.5)) +
  geom_errorbar(width = 0.2, position = position_dodge(width = 0.5)) +
  ylim(0, 0.75) +
  labs(
    title = "Set retrieval scores at rank 5 for Method A stratified by subject groups with 95% CI",
    x = "Subject Group",
    y = "Score"
  ) +
  facet_grid(rows = vars(metric)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

**Note:** You can also modify the previous bar-chart to include confidence intervals by
using the same approach as shown. Add the `compute_bootstrap_ci = TRUE` and
`n_bt` parameters to the `compute_set_retrieval_scores()` function calls and then
add the following lines to the ggplot code:

```r
geom_errorbar(
  mapping = aes(ymin = ci_lower, ymax = ci_upper),
  stat = "identity", 
  position = "dodge") +
```

## Bonus: Stratify by entity types of the vocabulary

In the above examples we stratified the results by document groups. It is also possible
to stratify by label groups, e.g. by entity types of the vocabulary. `CASIMiR` 
makes this easy by allowing to pass a data frame with label IDs and their 
corresponding groups to the `label_groups` argument of the
 `compute_set_retrieval_scores()` function.

```{r}
#| label: stratified-label-groups
#| cache: true

# optionally use parallel processing for faster computation
library(future)
plan(multicore)

res_at_5_by_entity_type_all_methods <- map_dfr(
  predictions,
  ~ compute_set_retrieval_scores(
    predicted = .x,
    gold_standard = gold_standard,
    label_groups = gnd_entity_types, # stratify by label entity types
    mode = "micro", # use micro-averaging for label_groups
    compute_bootstrap_ci = TRUE,
    n_bt = 100L,
    k = 5,
    progress = TRUE),
  .id = "Method"
)

plan(sequential) # reset to sequential processing

head(res_at_5_by_entity_type_all_methods)
```

**Note:** Here we set the `mode` parameter to `"micro"` to compute micro-averaged
scores across all documents for each entity type. This is often more meaningful when
stratifying by label groups, as some entity types may have very few associated labels.

```{r}
#| label: stratified-label-groups-plot
#| fig.width: 10
#| fig.height: 20
res_at_5_by_entity_type_all_methods  |>
  filter(metric == "f1", label_entitytype_eng != "NA") |>
  ggplot(aes(x = Method, y = value, fill = Method)) +
  ylim(0, 0.7) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(
    mapping = aes(ymin = ci_lower, ymax = ci_upper),
    stat = "identity", 
    position = "dodge"
  ) +
  facet_wrap(
    vars(label_entitytype_eng),
    ncol = 3,
    labeller = label_wrap_gen(width = 20)
  ) +
  labs(
    title = "F1-scores at rank 5 for all methods stratified by entity types",
    x = "Method",
    y = "F1-score  (micro averaged)"
  ) + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

Observe how the confidence intervals show extreme ranges for entity types with 
very few observed labels (e.g. conferences), and very narrow ranges for entity
 types with many labels, e.g. subject terms. 