---
title: "Workbook 4: Analysing the long-tail"
author: "Maximilian Kähler, DNB"
format: gfm
toc: true
editor: source
---

So far, we have mainly focused on evaluating the overall performance of
different methods for subject indexing. However, in practice, it is often
the case that some subjects are much more frequent than others. 


```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: FALSE
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.path = "figures/04_analysing-the-long-tail/"
  )

library(knitr)
library(tidyverse)
library(casimir)
options(casimir.ignore_inconsistencies = TRUE, digits = 3)
```

```{r}
#| label: read-data
#| echo: FALSE


gold_standard <- read_csv("../data/test-set_gold-standard.csv",
                          col_select = c("doc_id", "label_id"),
                          show_col_types = FALSE)

files <- list(
  "method-A" = "../data/test-set_predictions/method-A.csv",
  "method-B" = "../data/test-set_predictions/method-B.csv",
  "method-C" = "../data/test-set_predictions/method-C.csv",
  "method-D" = "../data/test-set_predictions/method-D.csv",
  "method-E" = "../data/test-set_predictions/method-E.csv",
  "method-F" = "../data/test-set_predictions/method-F.csv"
)
predictions <- files |> 
  map(read_csv, show_col_types = FALSE) 

# load training frequency distribution for GND labels
train_freqs <- read_csv("../data/training-frequency-distribution.csv",
                        show_col_types = FALSE) 
```

The following picture shows the frequency distribution of GND labels in the
training data.

```{r}
#| label: plot-frequency-distribution
 # compute binned frequency groups
freq_groups <- train_freqs  |>               
                mutate(freq_group = cut(x = label_freq,
                          breaks = c(0, 1, 10, 100, 1000, 10000, Inf),
                          labels = c("x == 0",
                                     "1 <= x < 10",
                                     "10 <= x < 100",
                                     "100 <= x < 1000",
                                     "1000 <= x < 10000",
                                     "10000 <= x"
                          ),
                          right = FALSE, include_lowest = TRUE))  |>
                  select(-label_freq)

freq_groups |> 
  filter(freq_group != "x == 0") |>
  group_by(freq_group) |> 
  summarise(n_labels = n()) |> 
  ggplot(aes(x = freq_group, y = n_labels)) + 
  geom_col(fill = "lightblue", color = "black") +
  labs(
    title = "Frequency distribution of GND labels in the training data",
    x = "x - Number of training documents per label",
    y = "Number of labels"
  )
```

The overwhelming majority of labels in the training data are very infrequent,
i.e., they occur less than 10 times. This is a common phenomenon in subject
indexing and is often referred to as the "long-tail" problem.
It is quite common that indexing methods perform well on frequent labels, but
struggle with infrequent ones. However, infrequent labels are often very important
for the precise description of documents, whereas frequent labels tend to be more
general and less informative. In the following, we will analyse how well the
different methods perform on labels of different frequencies.

## Analyse performance by label frequency, binned frequency groups

We already learned in workbook 02, how to stratify set retrieval metrics
by groups defined in the label space. We now use the same method, but
use the freq_groups derived from the training distribution above. 
```{r}
#| label: performance-by-frequency-group
#| cache: true

res_by_gnd_freq <- map_dfr(
  predictions,
  ~ compute_set_retrieval_scores(
    predicted = .x,
    gold_standard = gold_standard,
    mode = "micro",
    k = 5,
    label_groups = freq_groups
  ),  .id = "method"
)
```

We can visualize that in a plot:
```{r}
#| label: plot-performance-by-frequency-group
#| fig.width: 8
#| fig.height: 24

res_by_gnd_freq |> 
  filter(metric != "rprec", freq_group != "NA")  |>
  ggplot(aes(x = freq_group, y = value, fill = method)) + 
    geom_bar(stat = "identity", position = "dodge") +
    ylim(0, 1) +
    facet_grid(rows = vars(metric)) + 
    labs(
      title = "Precision, Recall and f1 by GND label frequency group",
      x = "GND label frequency group (in training data)",
      y = "Score (micro-averaged at k=5)"
    )
```

## Your turn

Based on the plot above, answer the following questions:

  * Which method performs best on infrequent labels (1-10 training instances)?
  * Which method performs best on frequent labels (>1000 training instances)?

What are your conclusions? Which method would you choose for a production system?

## Bonus: propensity scoring

