---
title: "Workbook 4: Analysing the long-tail"
author: "Maximilian Kähler, DNB"
format: gfm
toc: true
bibliography: ../slides/references.bib
editor: source
---

So far, we have mainly focused on evaluating the overall performance of
different methods for subject indexing. However, in practice, it is often
the case that some subjects are much more frequent than others. 


```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: FALSE
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.path = "figures/04_analysing-the-long-tail/"
  )

library(knitr)
library(tidyverse)
library(casimir)
options(casimir.ignore_inconsistencies = TRUE, digits = 3)
```

```{r}
#| label: read-data
#| echo: FALSE

source("../src/load_data.r")

# load training frequency distribution for GND labels
train_freqs <- read_csv("../data/training-frequency-distribution.csv",
                        show_col_types = FALSE)  |>
                # add total number of training documents
                mutate(n_docs = 1e7) 
```

The following picture shows the frequency distribution of GND labels in the
training data.

```{r}
#| label: plot-frequency-distribution
 # compute binned frequency groups
freq_groups <- train_freqs  |>               
                mutate(freq_group = cut(x = label_freq,
                          breaks = c(0, 1, 10, 100, 1000, 10000, Inf),
                          labels = c("x == 0",
                                     "1 <= x < 10",
                                     "10 <= x < 100",
                                     "100 <= x < 1000",
                                     "1000 <= x < 10000",
                                     "10000 <= x"
                          ),
                          right = FALSE, include_lowest = TRUE))  |>
                  select(-label_freq)

freq_groups |> 
  filter(freq_group != "x == 0") |>
  group_by(freq_group) |> 
  summarise(n_labels = n()) |> 
  ggplot(aes(x = freq_group, y = n_labels)) + 
  geom_col(fill = "lightblue", color = "black") +
  labs(
    title = "Frequency distribution of GND labels in the training data",
    x = "x - Number of training documents per label",
    y = "Number of labels"
  )
```

The overwhelming majority of labels in the training data are very infrequent,
i.e., they occur less than 10 times. This is a common phenomenon in subject
indexing and is often referred to as the "long-tail" problem.
It is quite common that indexing methods perform well on frequent labels, but
struggle with infrequent ones. However, infrequent labels are often very important
for the precise description of documents, whereas frequent labels tend to be more
general and less informative. In the following, we will analyse how well the
different methods perform on labels of different frequencies.

## Analyse performance by label frequency, binned frequency groups

We already learned in workbook 02 how to stratify set retrieval metrics
by groups defined in the label space. We now use the same method, but
use the freq_groups derived from the training distribution above. 
```{r}
#| label: performance-by-train-frequency-group
#| cache: true

res_by_train_freq <- map_dfr(
  predictions,
  ~ compute_set_retrieval_scores(
    predicted = .x,
    gold_standard = gold_standard,
    mode = "micro",
    k = 5,
    label_groups = freq_groups
  ),  .id = "method"
)
```

We can visualize that in a plot:
```{r}
#| label: plot-performance-by-frequency-group
#| fig.width: 8
#| fig.height: 24

res_by_train_freq |> 
  filter(metric != "rprec", freq_group != "NA")  |>
  ggplot(aes(x = freq_group, y = value, fill = method)) + 
    geom_bar(stat = "identity", position = "dodge") +
    ylim(0, 1) +
    facet_grid(rows = vars(metric)) + 
    labs(
      title = "Precision, Recall and f1 by GND label frequency group",
      x = "GND label frequency group (in training data)",
      y = "Score (micro-averaged at k=5)"
    )
```

## Your turn

Based on the plot above, answer the following questions:

  * Which method performs best on infrequent labels (1-10 training instances)?
  * Which method performs best on frequent labels (>1000 training instances)?



## Bonus: propensity scoring

Above analysis shows huge differences between the methods and strata of
frequency groups. However, it is difficult to summarize this information. 
An alternative to a stratified analysis along the frequency groups is to 
work with **weighted** labels. Going by the assumption that less frequent labels
hold more specific information and thus are of higher importance, we will
now define label weights that are inversly proportional to the label frequency. 

The precise formula for computing the weights is taken from [@PfastreXML]
and is based on the concept of propensity scoring. You can skip the details
of the formula in the box below, but feel free to expand it if you are interested.

::: {.panel-tabset}
#### More about propensity scores

<details>
<summary>Click to expand explanation of propensity scores</summary>

Can we assume that annotations in our gold standard are complete? Many subject
ontologies like the GND or LSCH have hundreds of thousands of labels, but
documents are usually only annotated with a handful of them. It is quite likely
that an annotator is not aware of all possible relevant labels. This is
especially true for untrained annotators. 
@PfastreXML argue the same and propose to model the incompleteness of the gold standard
by introducing propensity scores for each label $\lambda$. The propensity score
$p_{\lambda}$ models the probability that a relevant label $\lambda$ is actually
annotated, assuming it should be annotated (this is called a marginal probability).

$$p_{\lambda} = \frac{1}{1 + C\cdot exp(-A\cdot log(N_{\lambda} + B))}$$

Here, $N_{\lambda}$ is the number of training instances for label $\lambda$,
and $A$, $B$ are hyperparameters that can be tuned. $C$ is a constant that depends
on the total number of training instances $N$ (the entire training set size). 

The following plot shows the relationship between label frequency and
the corresponding label weight (inversly proportional to the propensity score).

```{r}

N_l <- 10^seq(0,6,0.5)

p_l <- function(N_l, N = 3e6, A = 0.55, B = 1.5)
 {
  C <- (log(N) - 1)*(B + 1)^A
  p_l <- 1/(1 + C*exp(-A*log(N_l + B)))
  
  return(p_l)
}


df <- expand.grid(
  N_l = N_l,
  N = c(1e7)
) %>% 
rowwise() %>% 
  mutate(p_l = p_l(N_l, N))

ggplot(df, aes(x = N_l, y = 1/p_l, color = factor(N))) + 
  geom_point() + 
  geom_smooth(se = FALSE) + 
  scale_x_log10() + 
  scale_color_discrete() + 
  labs(
    title = "Propensity scores",
    x = "Number of training instances per label (N_l)",
    y = "Label weight (inversely proportional to propensity score)",
    color = "Total number of training instances (N)"
  )
  
```

Rare labels receive much higher weights than frequent ones. 


</details>
:::

We will now use these label weights to compute weighted set retrieval metrics
that emphasize the performance on infrequent labels.
CASIMiR takes an additional argument `propensity_scored = TRUE` in the function
`compute_set_retrieval_scores()` (and also for the functions `compute_pr_auc()` and `compute_pr_curve()`). This will
require that you also pass a data frame with the label distribution in the training
data (as we have already loaded above in `train_freqs`).


```{r}
#| label: propensity-scored-metrics
#| cache: true

res_propensity <- map_dfr(
  predictions,
  ~ compute_set_retrieval_scores(
    predicted = .x,
    gold_standard = gold_standard,
    mode = "doc-avg",
    k = 5,
    propensity_scored = TRUE,
    label_distribution = train_freqs,
    rename_metrics = TRUE
  ),  .id = "method"
)

```

The following table shows the results:

```{r}
#| label: show-propensity-scored-metrics

res_propensity  |>
  filter(metric == "ps-f1@5")  |>
  kable(caption = "Propensity scored f1@5 scores by method")
```


### Conditional label weights

When we look at the errors an indexing algorithm makes we usually think of
false positive and false negatives. Do they mean the same to us?
If not, arguably, one might introduce what is called a "custom loss function"  

|                | Goldstandard yes | Goldstandard no |
|----------------|-----------------|-------------------|
| **Prediction yes**  | $C_{\text{tp}}\cdot \text{tp}$ | $C_{\text{fp}}\cdot\text{fp}$ |
| **Prediction no** | $C_{\text{fn}}\cdot\text{fn}$ | $C_{\text{tn}}\cdot\text{tn}$ |

with some custom weights assigned to each category
 $C_{\text{tp}}, C_{\text{fp}}, C_{\text{fn}}, C_{\text{tn}}$. 
 True negatives are never intresting in subject indexing, so lets ignore $C_{\text{tn}}$.
 Now, arguably, having a false positive prediction, i.e. having a document
 incorrectly labeled, should be punished irrespective of the label frequency. Having 
 a wrong frequent label is just as bad as having a wrong infrequent label. 
 However, when it comes to false negatives, i.e. missing a label that should have been
 assigned to a document, we might want to weight missing infrequent labels
 higher than missing frequent labels, still holding to the assumption that infrequent
labels are more specific and informative.
So we assume the following costs: 
$$C_{\text{tp}} = C_{\text{fn}} = w_{\lambda},$$
(with $w_{\lambda}$ the label weight derived from the propensity score as $1/p_{\lambda}$) and
$C_{\text{fp}} = \textit{const}$.

CASIMiR supports this by allowing to pass different a constant for false positives:
```{r}
#| label: propensity-scored-metrics-conditional
#| cache: true


res_propensity <- map_dfr(
  predictions,
  ~ compute_set_retrieval_scores(
    predicted = .x,
    gold_standard = gold_standard,
    mode = "doc-avg",
    k = 5,
    propensity_scored = TRUE,
    label_distribution = train_freqs,
    cost_fp_constant = "mean",
    rename_metrics = TRUE
  ),  .id = "method"
)

res_propensity  |>
  filter(metric == "ps-f1@5")  |>
  kable(caption = "Propensity scored f1@5 scores by method with constant false positive cost")
```

We can observe that `charming-cello` and `embracing-euphonium`, among the top 
notch in workbook 1, come closer to the other methods, whereas `bold-bassoon`
still stands out as the best method.

### Your turn

  * compare the results of the propensity scored metrics with the results
    stratified by frequency groups above. Do they tell a similar story?
  * use the argument `cost_fp_constant = ` to manipulate the false positive cost.
    Try different strategies and observe the changes in the results.

## References