---
title: "Workbook 7 (Bonus): Evaluating Graded Relevance Ratings"
author: "Maximilian Kähler, DNB"
format: gfm
toc: true
bibliography: ../slides/references.bib
editor: source
---

```{r}
#| label: setup
#| message: false
#| warning: false
#| echo: FALSE
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.path = "figures/07_evaluating-graded-relevance-ratings/"
  )

library(knitr)
library(tidyverse)
library(casimir)
options(casimir.ignore_inconsistencies = TRUE)
```

In this bonus workbook we will explore how to evaluate graded relevance ratings
using the CASIMiR package. Graded relevance ratings are useful when you want to
capture more nuanced judgments about the relevance of subject terms beyond a
simple binary relevant/ not relevant distinction. In binary relevance comparison,
we often find false postives that are not entirely wrong. Also manual indexers
can have some interrater disagreement, so that annotations are not always black and
white. Graded relevance ratings allow us to capture these nuances by assigning
different levels of relevance to machine based subject terms. This is the of 
evaluation that comes with the highes cost, as it requires subject experts to
manually rate the relevance of each suggested subject term.

## Datasets

In this workbook we use a different pair of data files for each method. For each
method we have a file of predicted subject terms along with graded relevance 
ratings assigned by subject experts and a gold standard file with binary relevance
labels, that also contain false negatives.

```{r}
#| label: read-graded-data
#| echo: FALSE

# load datasets as specified in r script
source("../src/load_data_graded_relevance.r")
source("../src/create_qual_table.r")

```

The graded relevance ratings are on an ordinal scale from 0 to 3:

  * 0: not relevant/ wrong
  * 1: slightly helpful
  * 2: helpful
  * 3: very helpful
  
Where helpful means that the subject term would be useful for a user to find the
document for retrieval purposes. To calculate metrics from these ordinal 
ratings, we map them to metric relevance levels $r$ between 0 and 1 as follows:

| Ordinal Relevance Level | Metric Relevance Level $r$ |
| ----------------------  | ---------------------- |
| very helpful | 1 |
| helpful      | 2/3 |
| slightly helpful | 1/3 |
| wrong  | 0 |

Let's take a look at how the data looks like for one of the methods,
e.g. "bold-bassoon".

```{r}
#| label: create-table

method_name <- "bold-bassoon"

preds <- predictions_w_relevance[[method_name]] 
gold <- gold_standards[[method_name]]
set.seed(681)
documents <- predictions_w_relevance[[method_name]] |> 
  distinct(doc_id) |>
  sample_n(3)

example_table <- create_qualitative_table_single_model(
    predicted = predictions_w_relevance[[method_name]],
    gold_standard = gold_standards[[method_name]],
    doc_id_list = documents,
    title_texts = select(doc_titles, doc_id, title = doc_title_eng),
    gnd = select(gnd, label_id, label_text = label_text_eng),
    graded_relevance = TRUE
) |> 
  group_by(doc_id, title_text)  |> 
  arrange(doc_id, gold, relevance) 

apply_styles_single_model_graded_rel(example_table)

```

### Your turn:

  * alter the seed to see more examples
  * try it out with other methods by changing the method name in the code above
  
## Graded Relevance Metrics

It is possible to generalise the binary relevance metrics Precision and Recall
to graded relevance ratings [@Kekalainen2002].

**Generalised Precision**  

$$
\mathrm{gPrec} := \frac{tp + \Delta_{rel}}{tp + fp}
$$

**Generalised Recall**  
$$
\mathrm{gRec} := \frac{tp + \Delta_{rel}}{tp + fn + \Delta_{rel}}
$$

with $\Delta_{rel} = \sum_{i \in false\ positives} r_i$  with metric relevance ratings $0 \leq r_i \leq 1$
  
## Computing Graded Relevance Metrics with CASIMiR

CASIMiR also provides functions to compute graded relevance metrics. 
Actually, you already know the `compute_set_retrieval_scores()` function from
workbook 2. It can also handle graded relevance ratings if you set the
`graded_relevance` argument to `TRUE`. This assumes that `predicted` contains
a `relevance` column with graded relevance ratings. 

The following example shows how to compute graded relevance metrics for the
"bold-bassoon" method.

```{r}
#| label: compute-graded-metrics

res <- compute_set_retrieval_scores(
  predicted = predictions_w_relevance[["bold-bassoon"]],
  gold_standard = gold_standards[["bold-bassoon"]],
  graded_relevance = TRUE,
  k = 5,
  rename_metrics = TRUE # adds prefix "g-" to metric names
) 

kable(
  res,
  caption = "Graded relevance metrics for bold-bassoon"
)

```

We can now iterate this over all methods like we did in workbook 2.

```{r}
#| label: compute-graded-metrics-multiple

res_all_methods <- map2_dfr(
  predictions_w_relevance,
  gold_standards,
  ~ compute_set_retrieval_scores(
    predicted = .x,
    gold_standard = .y,
    graded_relevance = TRUE,
    k = 5,
    rename_metrics = TRUE
  ),
  .id = "Method"
) |> 
  select(-support) |> 
  pivot_wider(
    names_from = metric,
    values_from = value
  ) 
```

**Note:** Workbook 2 was on a different dataset and here all methods were tested
on their own dataset. So the results are not directly comparable.

### Your turn

  * compare the results with the binary relevance metrics
  * can you compute a graded relevance precision recall curve?